---
title: "Fractionation analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in files and load packages


```{r}
library(tidyverse)
library(openxlsx)
library(gplots)
library(table1)

data <- read.xlsx("Input data/ProteinLevelRawRenamed.xlsx")  
LFQ_data <- data %>%
  filter(is.na(Reverse)) %>%
  filter(is.na(Potential.contaminant)) 

Twopeps <- LFQ_data %>%
  filter(Unique.peptides >= 2 )
##3843 proteins have at least 2 unique peptides


LFQ_data <- LFQ_data %>%
  select(Protein.IDs, LFQ_A1:LFQ_J7)

tmp <- LFQ_data %>%
  pivot_longer(-Protein.IDs, names_to = "Sample", values_to = "Quant") %>%
  filter(Quant >0) %>%
  group_by(Protein.IDs) %>%
  summarise(n = n()) %>%
  filter(n == 70)
##850 proteins are detected in all samples
  
write_csv(LFQ_data, "Input data/LFQ_quants_only.csv")

sample_key <- read.xlsx("Input data/SampleKey.xlsx")
sample_key_select <- sample_key %>%
  select(SampleID, Fancy_name)
table(sample_key$Condition)

sample_key <- sample_key %>%
   dplyr::select(Subject, Condition, Age, PMI) %>%
   unique()
 
 
 table1 <- table1(~ Age + PMI | Condition, data = sample_key)

```

##Protein yield differences

```{r}
yield <- read.xlsx("Input data/Protein yield table.xlsx")


yield_plot <- ggplot(data = yield, aes(x = Condition, y = `Final.Conc.(ug/uL)`)) + geom_boxplot() +  geom_point() +  facet_wrap(~Sample, scales = "free_y") + 
  ylab("Protein concentration (\U03BCg / \U03BCL)")

ggsave("Figures/FigS1_fracton_concentrations.pdf", plot = yield_plot, device = "pdf", width = 8.5, height = 6, units = "in", dpi = 300)


yield_ttest = yield %>%
  group_by(Sample) %>%
  summarise(n=n(),
            n_C=length(which(Condition=="Con")),
            n_AD=length(which(Condition=="AD")),
            pVal=t.test(`Final.Conc.(ug/uL)`[Condition=="Con"], 
                        `Final.Conc.(ug/uL)`[Condition=="AD"])$p.value) %>%
  mutate(padj = p.adjust(pVal, method = "BH")) %>%
  arrange(padj)

yield_ttest




```


## Plot fractions


```{r}
LFQ_data_long <- LFQ_data %>%
  pivot_longer(!Protein.IDs, names_to = "SampleID", values_to = "LFQ")

LFQ_data_long$LFQ[LFQ_data_long$LFQ == 0] <- NA

LFQ_data_long <- merge(sample_key_select, LFQ_data_long, by = "SampleID")

LFQ_dist <- ggplot(LFQ_data_long, aes(x = Fancy_name, y = log2(LFQ))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45, hjust=1)) + xlab("log2 LFQ") + ylab("Sample") +
  coord_flip()

ggsave("Figures/FigS2_LFQ_distributions.pdf", plot = LFQ_dist, device = "pdf", width = 8.5, height = 11, units = "in", dpi = 300)

LFQ_dist
```

No need to normalize any further.  This looks great.

##Substitute missing values with column minimum after filtering out any proteins found in fewer than 10 samples 

This is a pretty lax filter that allows for a protein to be detected only in one fraction.

```{r}

NA_summary <- LFQ_data_long %>%
  group_by(Protein.IDs) %>%
  summarize(n_NA = sum(is.na(LFQ))) 

ggplot(NA_summary, aes(x = n_NA)) + geom_histogram(binwidth = 2)

NA_summary_filtered <- NA_summary %>%
  filter(n_NA <= 60)

LFQ_data_long_filtered <- LFQ_data_long %>%
  filter(Protein.IDs %in% NA_summary_filtered$Protein.IDs)

LFQ_data_wide_filtered <- LFQ_data_long_filtered %>%
  pivot_wider(names_from = Protein.IDs, values_from = LFQ)

## Replace NAs with lowest LFQ value from that protein


LFQ_data_demos <- LFQ_data_wide_filtered[,1:2]
LFQ_data_matrix <- LFQ_data_wide_filtered[,3:ncol(LFQ_data_wide_filtered)]
LFQ_data_matrix <- LFQ_data_matrix %>%
   mutate_all(~ifelse(is.na(.x), min(.x, na.rm = TRUE), .x))

LFQ_data_wide_filtered <- cbind(LFQ_data_demos, LFQ_data_matrix)


```


```{r}
##Getting file ready for SVM
##Original data with columns for high missing data removed
SVM_LFQ_data <- LFQ_data %>%
  filter(Protein.IDs %in% NA_summary_filtered$Protein.IDs) %>%
  separate(Protein.IDs, sep = "\\|", into = c(NA, "UniprotID"), extra = "drop", remove = FALSE) %>%
  filter(!is.na(UniprotID)) %>%
  select(-Protein.IDs)

#Some duplicate IDs - sum to collapse
tmp <- which(duplicated(SVM_LFQ_data$UniprotID))
tmp <- SVM_LFQ_data[tmp,]
tmp <- tmp$UniprotID

duplicates_summed <- SVM_LFQ_data %>%
  filter(UniprotID %in% tmp) %>%
  arrange(UniprotID) %>%
  group_by(UniprotID) %>%
  summarise_all(.funs = sum)


SVM_LFQ_data <- SVM_LFQ_data %>%
  filter(!(UniprotID %in% tmp)) %>%
  bind_rows(duplicates_summed, .id = NULL) %>%
  column_to_rownames(var = "UniprotID")

tmp <- t(SVM_LFQ_data)
  
SVM_LFQ_data <- data.frame(tmp) %>%
  rownames_to_column() %>%
  separate(rowname, into = c(NA, "Code"), sep = "\\_") %>%
  separate(Code, into = c("Sample", "Fraction"), sep = 1)

write_csv(SVM_LFQ_data, "Input data/data_for_SVM.csv")


```


##PCA of all samples

```{r}

LFQ_data_logged <- log2(LFQ_data_wide_filtered[,3:ncol(LFQ_data_wide_filtered)])
rownames(LFQ_data_logged) <- LFQ_data_wide_filtered$Fancy_name

LFQ_PCA <- prcomp(LFQ_data_logged, center = TRUE, scale. = TRUE) 
plot(LFQ_PCA, type = "l")

summary(LFQ_PCA)
```

```{r}

PCA_df <- as.data.frame(LFQ_PCA$x)
PCA_df$group <- sapply( strsplit(as.character(row.names(PCA_df)), "_"), "[[", 2 )

head(PCA_df)

gp_col <- c("1" = "#d73027", "2" = "#fc8d59", "3" = "#fee090", "4" = "#fee090", "5" = "#e0f3f8", "6" = "#91bfdb", "7" = "#4575b4")

PCA_samples <- ggplot(PCA_df,aes(x=PC1,y=PC2)) + geom_point(aes(fill = group, size = 1, alpha = 0.8), colour="black",pch=21) + scale_fill_manual(values = gp_col) + theme_bw()

ggsave("Figures/Fig1B_sample_PCA.pdf", plot = PCA_samples, device = "pdf", width = 4.25, height = 3.66, units = "in", dpi = 300)


PCA_samples
```




## Heatmap - all proteins

```{r}

LFQ_data_to_plot <- LFQ_data_wide_filtered[,3:ncol(LFQ_data_wide_filtered)]
rownames(LFQ_data_to_plot) <- LFQ_data_wide_filtered$Fancy_name
LFQ_demos <- LFQ_data_wide_filtered[,1:2]
LFQ_demos$group <- sapply( strsplit(as.character(LFQ_demos$Fancy_name), "_"), "[[", 2 )

library(gplots)

gp_col_df <- data.frame(gp_col) %>%
  rownames_to_column() %>%
  rename(group = rowname) %>%
  left_join(LFQ_demos)

pdf(file="Figures/FigS3.pdf", width = 4.25, height = 20)

heatmap.2(as.matrix(log2(LFQ_data_to_plot)), trace = "none", 
          scale="col", 
          col=colorRampPalette(c("steelblue","white","darkred"))(50),
          RowSideColors = gp_col_df$gp_col[match(rownames(LFQ_data_to_plot), gp_col_df$Fancy_name)])

dev.off()

```

## ANOVA

```{r}
to_ANOVA <- cbind(LFQ_demos, LFQ_data_to_plot)
LFQ_quants <- as.matrix(to_ANOVA[,4:ncol(to_ANOVA)])

fit <- apply(LFQ_quants, 2, function(protein){ anova(lm(protein ~ as.factor(to_ANOVA$group)))})

anova <- data.frame(unlist(fit))
anova <- anova %>%
  rownames_to_column()

anova$protein <- sapply(strsplit(as.character(anova$rowname),'\\.'), "[", 1)
anova$outcome <- sapply(strsplit(as.character(anova$rowname),'\\.'), "[", 2)
 anova <- anova %>%
   select(-rowname) %>%
   spread(outcome, unlist.fit.) %>%
   select(-`F value2`, -`Pr(>F)2`)
anova$padj <- p.adjust(anova$`Pr(>F)1`, method = "BH")
anova <- arrange(anova, padj)
anova$Significant <- ifelse(anova$padj < 0.05, "Yes", "No")

table(anova$Significant)

sig_proteins <- anova %>%
  filter(Significant == "Yes")

sig_proteins_vector <- c(sig_proteins$protein) 

ANOVA_results <- anova


```   


## Heatmap of sig. proteins

```{r}

LFQ_data_to_plot_filtered <- LFQ_data_wide_filtered[,3:ncol(LFQ_data_wide_filtered)]
LFQ_data_to_plot_filtered <- LFQ_data_to_plot_filtered %>%
  select(all_of(sig_proteins_vector))

rownames(LFQ_data_to_plot_filtered) <- LFQ_data_wide_filtered$Fancy_name


pdf(file="Figures/Fig1D.pdf", width = 4.25, height = 20)

heatmap.2(as.matrix(log2(LFQ_data_to_plot_filtered)), trace = "none", 
          scale="col", 
          col=colorRampPalette(c("steelblue","white","darkred"))(50),
          RowSideColors = gp_col_df$gp_col[match(rownames(LFQ_data_to_plot_filtered), gp_col_df$Fancy_name)])

dev.off()



```


## Tukey testing - pairwise comparisons

```{r}

fit_for_Tukey <- apply(LFQ_quants, 2, function(protein){ aov(lm(protein ~ as.factor(to_ANOVA$group)))})

#fit_for_Tukey <- apply(LFQ_quants, 2, function(variable){ aov(lm(variable ~ Dx$Project.diagnosis))})

Tukey_output <- lapply(fit_for_Tukey, TukeyHSD)


for(i in seq_along(Tukey_output)) {
  Tukey_output[[i]] <- as.data.frame(Tukey_output[[i]]$`as.factor(to_ANOVA$group)`)
  Tukey_output[[i]] <- rownames_to_column(Tukey_output[[i]])
}
#Collapse list into one large dataframe for all_data
df_all_tukey <- map_dfr(Tukey_output, as.data.frame, .id = "id")
df_all_tukey$Significant <- ifelse(df_all_tukey$`p adj` < 0.05, "Yes", "No")


#this includes proteins that didn't survive multiple test corrections at the anova level
df_all_tukey_filtered <- df_all_tukey %>%
  filter(id %in% sig_proteins_vector)
  
  
sig_comparisons <- df_all_tukey_filtered %>%
  filter(Significant == "Yes")

head(sig_comparisons)

table(sig_comparisons$rowname)

wide_sig_comparisons <- sig_comparisons %>%
  select(id, rowname, `p adj`) %>%
  pivot_wider(names_from = rowname, values_from = `p adj`, values_fill = NA) %>%
  select(id, `2-1`, `3-1`, `4-1`, `5-1`, `6-1`, `7-1`, `3-2`, `4-2`, `5-2`, `6-2`, `7-2`, `4-3`, `5-3`, `6-3`, `7-3`, `5-4`, `6-4`, `7-4`, `6-5`, `7-5`, `7-6`)

write_csv(sig_comparisons, "tmp/sig_proteins.csv")

n_sig <- data.frame(table(sig_comparisons$rowname))

ANOVA_results <- merge(ANOVA_results, wide_sig_comparisons, by.x = "protein", by.y = "id", all.x = TRUE)
write_csv(ANOVA_results, "Supp tables/S3_ANOVA_results.csv")
```

#Unique fraction proteins?

```{r}
list_gen <- function(col){
  wide_sig_comparisons %>%
    filter(!is.na(col)) %>%
    pull(id)
}



sig_list <- apply(wide_sig_comparisons[,-1], 2, list_gen)

library(UpSetR)

pdf(file="Figures/Fig1B_upset.pdf", width = 8.5, height = 6)
upset(fromList(sig_list), order.by = "freq", nsets = 10, nintersects = 20)
dev.off()


```

## Where are the marker proteins? 

```{r}

Uniprot_LFQ_data_long_filtered <- LFQ_data_long_filtered %>%
  separate(Protein.IDs, sep = "\\|", into = c(NA, "UniprotID"), extra = "drop", remove = FALSE)



marker_list <- read_csv("Input data/gold_markers.csv") %>%
  rename(Marker = ProteinID)


markers <- merge(Uniprot_LFQ_data_long_filtered, marker_list, by.x = "UniprotID", by.y = "Marker", all.x = TRUE)




```


```{r}

Fraction_data_logged <- t(LFQ_data_logged)

tmpA <- grep("P14136", rownames(Fraction_data_logged))
tmpB <- grep("P68871", rownames(Fraction_data_logged))

Fraction_data_logged <- Fraction_data_logged[-c(tmpA,tmpB),]   

tmp <- data.frame(rownames(Fraction_data_logged)) 
tmp <- tmp %>%
  separate(rownames.Fraction_data_logged., sep = "\\|", into = c(NA, "UniprotID"), extra = "drop", remove = TRUE) %>%
  pull(UniprotID)
rownames(Fraction_data_logged) <- tmp


Fraction_PCA <- prcomp(Fraction_data_logged, center = TRUE, scale. = TRUE) 
plot(Fraction_PCA, type = "l")

summary(Fraction_PCA)
```

```{r}

Fraction_PCA_df <- as.data.frame(Fraction_PCA$x)
Fraction_PCA_df <- Fraction_PCA_df %>%
  rownames_to_column() %>%
  rename(Uniprot = rowname)
Fraction_PCA_df <- merge(marker_list, Fraction_PCA_df, by.x = "Marker", by.y = "Uniprot", all.y = TRUE)


PCA_1_2 <- Fraction_PCA_df %>%
  filter(!is.na(Location)) %>%
  ggplot(aes(x=PC1,y=PC2)) + geom_point(aes(fill = Location, size = 1, alpha = 0.8), color ="black", pch =  21) + 
  scale_fill_brewer(palette = "Set3") + theme_bw()

ggsave("Figures/Fig2A_PCA12.pdf", plot = PCA_1_2, device = "pdf", width = 6, height = 3.6, units = "in", dpi = 300)

facet_1_2 <- Fraction_PCA_df %>%
  filter(!is.na(Location)) %>%
  ggplot(aes(x=PC1,y=PC2)) + geom_point(aes(fill = Location, size = 1, alpha = 0.8), color ="black", pch =  21) + 
  scale_fill_brewer(palette = "Set3") + theme_bw() + facet_wrap(~Location)

ggsave("Figures/FigS4A_PCA12_facet.pdf", plot = facet_1_2, device = "pdf", width = 10, height = 5.75, units = "in", dpi = 300)

PCA_2_3 <- Fraction_PCA_df %>%
  filter(!is.na(Location)) %>%
  ggplot(aes(x=PC2,y=PC3)) + geom_point(aes(fill = Location, size = 1, alpha = 0.8), color ="black", pch =  21) + 
  scale_fill_brewer(palette = "Set3") + theme_bw()

ggsave("Figures/Fig2B_PCA23.pdf", plot = PCA_2_3, device = "pdf", width = 6, height = 3.6, units = "in", dpi = 300)

facet_2_3 <- Fraction_PCA_df %>%
  filter(!is.na(Location)) %>%
  ggplot(aes(x=PC2,y=PC3)) + geom_point(aes(fill = Location, size = 1, alpha = 0.8), color ="black", pch =  21) + 
  scale_fill_brewer(palette = "Set3") + theme_bw() + facet_wrap(~Location)

ggsave("Figures/FigS4B_PCA23_facet.pdf", plot = facet_2_3, device = "pdf", width = 10, height = 5.75, units = "in", dpi = 300)



PC_boxplots <- Fraction_PCA_df %>%
  filter(!is.na(Location)) %>%
  pivot_longer(-c(Marker, Location), names_to = "PC", values_to = "Loading") %>%
  mutate(PC = gsub("PC", "", PC)) %>%
  mutate(PC = as.numeric(PC)) %>%
  filter( PC <5) %>%
  ggplot(aes (x= Location, y = Loading, fill = Location)) + 
  geom_boxplot() + facet_wrap(~PC, scales = "free", ncol = 2) + 
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + scale_fill_brewer(palette = "Set3")

ggsave("Figures/Fig2C_PCA_boxplots.pdf", plot = PC_boxplots, device = "pdf", width = 10, height = 5.75, units = "in", dpi = 300)

PCA_3_4 <- Fraction_PCA_df %>%
  filter(!is.na(Location)) %>%
  ggplot(aes(x=PC3,y=PC4)) + geom_point(aes(fill = Location, size = 1, alpha = 0.8), color ="black", pch =  21) + 
  scale_fill_brewer(palette = "Set3") + theme_bw()
```

## Marker average distribution

```{r}
controls <- c("A", "C", "E", "G", "I")
control_data <- SVM_LFQ_data %>%
  filter(Sample %in% controls) %>%
  pivot_longer(-c(Sample, Fraction), names_to = "ProteinID", values_to = "LFQ")
  
summarized_control_data <- control_data %>%
  group_by(ProteinID, Fraction) %>%
  summarize(meanLFQ = mean(LFQ))

protein_totals <- summarized_control_data %>%
  group_by(ProteinID) %>%
  summarize(total = sum(meanLFQ))

summarized_control_data <- merge(summarized_control_data, protein_totals, by = "ProteinID")

wide_summarized_control_data <- summarized_control_data %>%
  mutate(proportion = (meanLFQ/total)*100) %>%
  dplyr::select(-meanLFQ) %>%
  pivot_wider(names_from = Fraction, values_from = proportion)

wide_markers <- merge(wide_summarized_control_data, marker_list, by.x = "ProteinID", by.y = "Marker", all.x = FALSE)
long_markers <- wide_markers %>%
  pivot_longer(-c(ProteinID, Location, total), names_to = "Fraction", values_to = "Proportion")


Fig2D <- long_markers %>%
  group_by(Location, Fraction) %>%
  summarise(mean_prop = mean(Proportion)) %>%
  ggplot(aes(x = Fraction, y = mean_prop, group = Location)) + geom_line(aes(color = Location)) + 
    scale_colour_brewer(palette = "Set3") + theme_bw() + ylab("Mean proportion of LFQ intensity")



ggsave("Figures/Fig2D_marker_distribution.pdf", plot = Fig2D, device = "pdf", width = 6, height = 3.6, units = "in", dpi = 300)

Fig2D

```







