---
title: "SVM"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(openxlsx)
```

## Read in data


```{r}
for_SVM <- read_csv("Input data/data_for_SVM.csv")
sample_key <- read.xlsx("Input data/SampleKey.xlsx")
sample_key <- sample_key %>%
  dplyr::select(Subject, Condition, Age, PMI) %>%
  unique()
marker_list1 <- read_csv("Input data/gold_markers.csv")
marker_list1 <- unique(marker_list1)


table(marker_list1$Location)
```

## Test SVM with all markers

```{r}


SampleA <- for_SVM %>%
  filter(Sample == "A") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleA <- t(SampleA)
SampleA <- data.frame(SampleA) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)



SampleAmerged <- merge(SampleA, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)





SampleAtrain <- SampleAmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  column_to_rownames(var = "UniprotID") 
 

```

Not very balanced categories, so this might be a bit tough.




## Run a test SVM

```{r}

down_SampleA <- downSample(x = SampleAtrain, y = as.factor(SampleAtrain$Location))
down_SampleA <- down_SampleA %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)

set.seed(2709)
svM_A <- train(Location ~ ., data = down_SampleA, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_A$bestTune

svM_A

conf <- confusionMatrix(svM_A)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
Endolysosome <- (conf[3,3] / sum(conf[,3])) *100
`Endoplasmic Reticulum` <- (conf[4,4] / sum(conf[,4])) *100
`Golgi Apparatus` <- (conf[5,5] / sum(conf[,5])) *100
Mitochondria <- (conf[6,6] / sum(conf[,6])) *100
Nucleus <- (conf[7,7] / sum(conf[,7])) *100
Peroxisome <- (conf[8,8] / sum(conf[,8])) *100
`Plasma membrane` <- (conf[9,9] / sum(conf[,9])) *100
Postsynapse <- (conf[10,10] / sum(conf[,10])) *100
Presynapse <- (conf[11,11] / sum(conf[,11])) *100

conf
```

This is pretty bad given it has four proteins per category.  Filter out these low n categories.

```{r}

tmp <- data.frame(table(SampleAmerged$Location)) %>%
  filter(Freq >=35)

SampleAtrain <- SampleAmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Rerun SVM

```{r}
set.seed(2409)
down_SampleA <- downSample(x = SampleAtrain, y = as.factor(SampleAtrain$Location))
down_SampleA <- down_SampleA %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)

set.seed(2709)
svM_A <- train(Location ~ ., data = down_SampleA, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_A$bestTune

svM_A

conf <- confusionMatrix(svM_A)
conf <- conf$table

Cytoplasm2 <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton2 <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum2` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria2 <- (conf[4,4] / sum(conf[,4])) *100
Nucleus2 <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane2` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse2<- (conf[7,7] / sum(conf[,7])) *100
Presynapse2 <- (conf[8,8] / sum(conf[,8])) *100
results <- data.frame(Organelle = c("Cytoplasm", "Cytoskeleton", "Endoplasmic Reticulum", "Mitochondria", "Nucleus", "Plasma membrane", "Postsynapse", "Presynapse"), SampleA = c(Cytoplasm2, Cytoskeleton2, `Endoplasmic Reticulum2`, Mitochondria2, Nucleus2, `Plasma membrane2`, Postsynapse2, Presynapse2))

conf
```

##Sample C - reduced training set only

```{r}
SampleC <- for_SVM %>%
  filter(Sample == "C") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleC <- t(SampleC)
SampleC <- data.frame(SampleC) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)



SampleCmerged <- merge(SampleC, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleCtrain <- SampleCmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample C SVM


```{r}

down_SampleC <- downSample(x = SampleCtrain, y = as.factor(SampleCtrain$Location))
down_SampleC <- down_SampleC %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)

set.seed(2709)
svM_C <- train(Location ~ ., data = down_SampleC, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_C$bestTune

svM_C

conf <- confusionMatrix(svM_C)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleC <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)

results
```


##Sample E - reduced training set only

```{r}
SampleE <- for_SVM %>%
  filter(Sample == "E") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleE <- t(SampleE)
SampleE <- data.frame(SampleE) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)


SampleEmerged <- merge(SampleE, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleEtrain <- SampleEmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample C SVM


```{r}

down_SampleE <- downSample(x = SampleEtrain, y = as.factor(SampleEtrain$Location))
down_SampleE <- down_SampleE %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)


svM_E <- train(Location ~ ., data = down_SampleE, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_E$bestTune

svM_E

conf <- confusionMatrix(svM_E)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleE <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)

results
```

##Sample G - reduced training set only

```{r}
SampleG <- for_SVM %>%
  filter(Sample == "G") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleG <- t(SampleG)
SampleG <- data.frame(SampleG) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)


SampleGmerged <- merge(SampleG, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleGtrain <- SampleGmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample C SVM


```{r}

down_SampleG <- downSample(x = SampleGtrain, y = as.factor(SampleGtrain$Location))
down_SampleG <- down_SampleG %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)


svM_G <- train(Location ~ ., data = down_SampleG, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_G$bestTune

svM_G

conf <- confusionMatrix(svM_G)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleG <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)

results
```


##Sample I - reduced training set only

```{r}
SampleI <- for_SVM %>%
  filter(Sample == "I") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleI <- t(SampleI)
SampleI <- data.frame(SampleI) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)


SampleImerged <- merge(SampleI, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleItrain <- SampleImerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample I SVM


```{r}

down_SampleI <- downSample(x = SampleItrain, y = as.factor(SampleItrain$Location))
down_SampleI <- down_SampleI %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)


svM_I <- train(Location ~ ., data = down_SampleI, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_I$bestTune

svM_I

conf <- confusionMatrix(svM_I)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleI <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)


Con_train_results <- results %>%
  mutate(mean_perf = (SampleA + SampleC + SampleE + SampleG + SampleI)/5)
results

write_csv(Con_train_results, "tmp/control_classification_results.csv")


```

##  Compare with PMI

```{r}
PMI <- Con_train_results %>%
  dplyr::select(-mean_perf) %>%
  pivot_longer(-Organelle, names_to = "Sample", values_to = "performance") %>%
  separate(Sample, sep = 6, into = c(NA, "Subject"), remove = FALSE) %>%
  left_join(sample_key) 

ggplot(PMI, aes(x = as.factor(PMI), y = performance)) + geom_boxplot()
ggplot(PMI, aes(x = as.factor(PMI), y = performance, group = Organelle)) + geom_line(aes(color = Organelle))


```

No clear effect of increasing PMI between 8 and 23 hours - what happens earlier?

##Prediction for the rest of the proteins

```{r}
##Filter out training proteins

SampleAtest <- SampleAmerged %>%
  anti_join(down_SampleA) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleAtest$predicted.locations_A <- predict(svM_A, newdata = SampleAtest)
SampleAtest <- SampleAtest %>%
  rownames_to_column()

table(SampleAtest$predicted.locations_A)

write_csv(SampleAtest, "tmp/SampleA_test.csv")
```

##Sample C

```{r}
##Filter out training proteins

SampleCtest <- SampleCmerged %>%
  anti_join(down_SampleC) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleCtest$predicted.locations_C <- predict(svM_C, newdata = SampleCtest)
SampleCtest <- SampleCtest %>%
  rownames_to_column()

```

##Sample E

```{r}
##Filter out training proteins

SampleEtest <- SampleEmerged %>%
  anti_join(down_SampleE) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleEtest$predicted.locations_E <- predict(svM_E, newdata = SampleEtest)
SampleEtest <- SampleEtest %>%
  rownames_to_column()

```

##Sample G

```{r}
##Filter out training proteins

SampleGtest <- SampleGmerged %>%
  anti_join(down_SampleG) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleGtest$predicted.locations_G <- predict(svM_G, newdata = SampleGtest)
SampleGtest <- SampleGtest %>%
  rownames_to_column()

```


##Sample I

```{r}
##Filter out training proteins

SampleItest <- SampleImerged %>%
  anti_join(down_SampleI) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleItest$predicted.locations_I <- predict(svM_I, newdata = SampleItest)
SampleItest <- SampleItest %>%
  rownames_to_column()

```


##Results table

```{r}
tmpA <- SampleAtest %>%
  dplyr::select(rowname, predicted.locations_A)
tmpC <- SampleCtest %>%
  dplyr::select(rowname, predicted.locations_C)
tmpE <- SampleEtest %>%
  dplyr::select(rowname, predicted.locations_E)
tmpG <- SampleGtest %>%
  dplyr::select(rowname, predicted.locations_G)
tmpI <- SampleItest %>%
  dplyr::select(rowname, predicted.locations_I)




test_results <- merge(tmpA, tmpC, by = "rowname", all = TRUE)
test_results <- merge(test_results, tmpE, by = "rowname", all = TRUE)
test_results <- merge(test_results, tmpG, by = "rowname", all = TRUE)
test_results <- merge(test_results, tmpI, by = "rowname", all = TRUE)

test_results$N <- apply(!is.na(test_results[,2:6]), 1, sum)
n_group <- test_results %>%
  dplyr::select(rowname, N)

tmp <- test_results %>%
  pivot_longer(-c(rowname, N), names_to = "Sample", values_to = "Location") 

tmp <- data.frame(table(tmp$rowname, tmp$Location))
results_consensus <- tmp %>%
  group_by(Var1) %>%
  slice_max(Freq)


results_consensus <- merge(results_consensus, n_group, by.x = "Var1", by.y = "rowname")
results_consensus <- results_consensus %>%
  mutate(prop = (Freq/N)*100)

clear_location <- results_consensus %>%
  filter(prop > 50) %>%
  mutate(Confidence = ifelse(prop > 70, "High", "Medium")) %>%
  dplyr::select(Var1, Var2, Confidence) %>%
  dplyr::rename(Location = Var2)


ties <- results_consensus %>%
  filter(!(Var1 %in% clear_location$Var1)) %>%
  arrange(Var1) 


fourty_percent <- ties %>%
  filter(prop >= 40) %>%
  group_by(Var1) %>%
  mutate(id = row_number()) %>%
  dplyr::select(Var1, id, Var2) %>%
  arrange(Var2) %>%
  pivot_wider(names_from = id, values_from = Var2) %>%
  dplyr::rename(Location_1 = `1`, Location_2 = `2`) %>%
  unite("Split_location", Location_1:Location_2, remove = FALSE) 

fourty_percent$Split_location <- gsub("_NA", "", fourty_percent$Split_location)
low_confidence <- fourty_percent %>%
  mutate(Confidence = "Low") %>%
  dplyr::select(Var1, Split_location, Confidence) %>%
  dplyr::rename(Location = Split_location)

remaining <- results_consensus %>%
  filter(!(Var1 %in% clear_location$Var1)) %>%
  filter(!(Var1 %in% fourty_percent$Var1)) %>%
  mutate(Location = NA) %>%
  mutate(Confidence = "None") %>%
  dplyr::select(Var1, Location, Confidence) %>%
  unique()
 
##Change confidence from high group to high and Medium
final_results <- rbind(clear_location, low_confidence, remaining)
final_results <- final_results %>%
  arrange(Confidence)

tmp <- data.frame(table(final_results$Location))

table(final_results$Location, final_results$Confidence)


head(final_results)

```


## Ontology with TOPGO

Needs ID conversion from Uniprot


```{r}

library(org.Hs.eg.db)

uniKeys <- keys(org.Hs.eg.db, keytype="UNIPROT")
cols <- c("SYMBOL")
conversion_Ref <- select(org.Hs.eg.db, keys=uniKeys, columns=cols, keytype="UNIPROT")



```

## ER annotated enrichment

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Endoplasmic Reticulum" & Confidence == "High", 0.01, 1))

library(topGO)
#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_ER <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_ER$SVM_fraction <- "ER"




```

## Plasma Membrane annotated

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Plasma membrane" & Confidence == "High", 0.01, 1))


#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_Plas_memb <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_Plas_memb$SVM_fraction <- "Plasma membrane"




```


## Cytoplasm annotated

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Cytoplasm" & Confidence == "High", 0.01, 1))


#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_cytoplasm <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_cytoplasm$SVM_fraction <- "Cytoplasm"




```

## Cytoskeleton annotated

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Cytoskeleton" & Confidence == "High", 0.01, 1))


#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_cytoskel <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_cytoskel$SVM_fraction <- "Cytoskeleton"




```

## Nucleus annotated

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Nucleus" & Confidence == "High", 0.01, 1))


#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_nuc <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_nuc$SVM_fraction <- "Nucleus"




```


## Mitochondria annotated

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Mitochondria" & Confidence == "High", 0.01, 1))


#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_mito <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_mito$SVM_fraction <- "Mitochondria"




```

## Postsynapse annotated

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Postsynapse" & Confidence == "High", 0.01, 1))


#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_post <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_post$SVM_fraction <- "Postsynapse"




```


## Presynapse annotated

```{r}

results <- final_results %>%
  dplyr::rename(UNIPROT = Var1) %>%
  left_join(conversion_Ref)%>%
  mutate(sig_val = ifelse(Location == "Presynapse" & Confidence == "High", 0.01, 1))


#Convert to vector
allGenes_input <- setNames(as.numeric(results$sig_val), results$SYMBOL)

#Build function to dplyr::select Geneset (significant genes)
selFun <- function(x){return(x < 0.05)}

#Build GO data object - for biological process
GO_data_CC <- new("topGOdata", 
               ontology = "CC",
               allGenes = allGenes_input,
               geneSel = selFun,
               annot = annFUN.org,
               mapping = "org.Hs.eg.db",
               ID = "symbol"
)

CC_result <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")


CC_results_pre <- GenTable(GO_data_CC, classicFisher = CC_result, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 5)
CC_results_pre$SVM_fraction <- "Presynapse"


All_CC_results <- rbind(CC_results_cytoplasm, CC_results_cytoskel, CC_results_ER, CC_results_mito, CC_results_nuc, CC_results_Plas_memb, CC_results_post, CC_results_pre)

write_csv(All_CC_results, "Supp tables/Table1_high_confidence_GO_enrichment.csv")

All_CC_results
```

## AD Samples

```{r}


SampleB <- for_SVM %>%
  dplyr::filter(Sample == "B") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleB <- t(SampleB)
SampleB <- data.frame(SampleB) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)



SampleBmerged <- merge(SampleB, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

```


```{r}

tmp <- data.frame(table(SampleBmerged$Location)) %>%
  filter(Freq >=35)

SampleBtrain <- SampleBmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Rerun SVM

```{r}
set.seed(2409)
down_SampleB <- downSample(x = SampleBtrain, y = as.factor(SampleBtrain$Location))
down_SampleB <- down_SampleB %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)

set.seed(2709)
svM_B <- train(Location ~ ., data = down_SampleB, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_B$bestTune

svM_B

conf <- confusionMatrix(svM_B)
conf <- conf$table

Cytoplasm2 <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton2 <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum2` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria2 <- (conf[4,4] / sum(conf[,4])) *100
Nucleus2 <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane2` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse2<- (conf[7,7] / sum(conf[,7])) *100
Presynapse2 <- (conf[8,8] / sum(conf[,8])) *100
results <- data.frame(Organelle = c("Cytoplasm", "Cytoskeleton", "Endoplasmic Reticulum", "Mitochondria", "Nucleus", "Plasma membrane", "Postsynapse", "Presynapse"), SampleB = c(Cytoplasm2, Cytoskeleton2, `Endoplasmic Reticulum2`, Mitochondria2, Nucleus2, `Plasma membrane2`, Postsynapse2, Presynapse2))

conf
```

##Sample D - reduced training set only

```{r}
SampleD <- for_SVM %>%
  filter(Sample == "D") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleD <- t(SampleD)
SampleD <- data.frame(SampleD) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)



SampleDmerged <- merge(SampleD, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleDtrain <- SampleDmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample D SVM


```{r}

down_SampleD <- downSample(x = SampleDtrain, y = as.factor(SampleDtrain$Location))
down_SampleD <- down_SampleD %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)

set.seed(2709)
svM_D <- train(Location ~ ., data = down_SampleD, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_D$bestTune

svM_D

conf <- confusionMatrix(svM_D)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleD <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)

results
```


##Sample F - reduced training set only

```{r}
SampleF <- for_SVM %>%
  filter(Sample == "F") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleF <- t(SampleF)
SampleF <- data.frame(SampleF) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)


SampleFmerged <- merge(SampleF, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleFtrain <- SampleFmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample F SVM


```{r}

down_SampleF <- downSample(x = SampleFtrain, y = as.factor(SampleFtrain$Location))
down_SampleF <- down_SampleF %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)


svM_F <- train(Location ~ ., data = down_SampleF, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_F$bestTune

svM_F

conf <- confusionMatrix(svM_F)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleF <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)

results
```

##Sample H - reduced training set only

```{r}
SampleH <- for_SVM %>%
  filter(Sample == "H") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleH <- t(SampleH)
SampleH <- data.frame(SampleH) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)


SampleHmerged <- merge(SampleH, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleHtrain <- SampleHmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample H SVM


```{r}

down_SampleH <- downSample(x = SampleHtrain, y = as.factor(SampleHtrain$Location))
down_SampleH <- down_SampleH %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)


svM_H <- train(Location ~ ., data = down_SampleH, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_H$bestTune

svM_H

conf <- confusionMatrix(svM_H)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleH <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)

results
```


##Sample J - reduced training set only

```{r}
SampleJ <- for_SVM %>%
  filter(Sample == "J") %>%
  dplyr::select(-Sample) %>%
  column_to_rownames(var = "Fraction")
SampleJ <- t(SampleJ)
SampleJ <- data.frame(SampleJ) %>%
  rownames_to_column() %>%
  dplyr::rename(UniprotID = rowname)


SampleJmerged <- merge(SampleJ, marker_list1, by.x = "UniprotID",by.y = "ProteinID",  all.x = TRUE)

SampleJtrain <- SampleJmerged %>%
  dplyr::filter(!is.na(Location)) %>%
  dplyr::filter(Location %in% tmp$Var1) %>%
  column_to_rownames(var = "UniprotID") 

```


## Sample J SVM


```{r}

down_SampleJ <- downSample(x = SampleJtrain, y = as.factor(SampleJtrain$Location))
down_SampleJ <- down_SampleJ %>%
  dplyr::select(-Class)

trControl <- trainControl("cv", number = 10, returnData = TRUE)


svM_J <- train(Location ~ ., data = down_SampleJ, method = "svmRadial", trControl = trControl, preProcess = c("center","scale"), tuneLength = 10)

svM_J$bestTune

svM_J

conf <- confusionMatrix(svM_J)
conf <- conf$table

Cytoplasm <- (conf[1,1] / sum(conf[,1])) *100
Cytoskeleton <- (conf[2,2] / sum(conf[,2])) *100
`Endoplasmic Reticulum` <- (conf[3,3] / sum(conf[,3])) *100
Mitochondria <- (conf[4,4] / sum(conf[,4])) *100
Nucleus <- (conf[5,5] / sum(conf[,5])) *100
`Plasma membrane` <- (conf[6,6] / sum(conf[,6])) *100
Postsynapse<- (conf[7,7] / sum(conf[,7])) *100
Presynapse <- (conf[8,8] / sum(conf[,8])) *100

results$SampleJ <- c(Cytoplasm, Cytoskeleton, `Endoplasmic Reticulum`, Mitochondria, Nucleus, `Plasma membrane`, Postsynapse, Presynapse)


AD_train_results <- results %>%
  mutate(mean_perf = (SampleB + SampleD + SampleF + SampleH + SampleJ)/5)
AD_train_results

write_csv(AD_train_results, "tmp/control_classification_results_AD.csv")


```

##  Compare with PMI

```{r}
Con_results_long <- Con_train_results %>%
  dplyr::select(-mean_perf) %>%
  pivot_longer(-Organelle, names_to = "Sample", values_to = "performance") %>%
  separate(Sample, sep = 6, into = c(NA, "Subject"), remove = FALSE)

All_results_long <- AD_train_results %>%
  dplyr::select(-mean_perf) %>%
  pivot_longer(-Organelle, names_to = "Sample", values_to = "performance") %>%
  separate(Sample, sep = 6, into = c(NA, "Subject"), remove = FALSE) %>%
  bind_rows(Con_results_long) %>%
  left_join(sample_key)

SVM_train_condition <- ggplot(All_results_long, aes(x=Organelle, y = performance, fill = Condition)) + 
  geom_boxplot() + geom_point(position = position_dodge(0.75)) + theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) 

ggsave("Figures/FigS5A_SVM_performance.pdf", plot = SVM_train_condition, device = "pdf", width = 8.5, height = 4, units = "in", dpi = 300)

SVM_train_condition
```




##Effect of PMI

```{r}
dummy_df <- data.frame(Organelle = c("Endolysosome", "Golgi Apparatus", "Peroxisome"), PMI = c(NA, NA, NA), performance = c(NA, NA, NA))

PMI <- All_results_long %>%
  group_by(Organelle, PMI) %>%
  summarise(performance = mean(performance)) %>%
  bind_rows(dummy_df) %>%
  ggplot(aes(x = as.factor(PMI), y = performance, group = Organelle, color = Organelle)) + geom_line() +
  xlab("PMI (hours)") + scale_color_brewer(palette = "Set3") + theme_bw()

ggsave("Figures/FigS5B_PMI_performance.pdf", plot = PMI, device = "pdf", width = 8.5, height = 4, units = "in", dpi = 300)

PMI
```


##lm to check

```{r}

lm_condition <- lm(performance ~ Condition * Organelle + Age + PMI, data = All_results_long)
lm_condition <- lm(performance ~ Condition * Organelle, data = All_results_long)

summary(lm_condition)


all_results_summary <- All_results_long %>%
  group_by(Organelle) %>%
  summarise(mean = mean(performance))

```

There are clear differences between Organelle in performance of the SVM on training data, but this does not appear to be affected by PMI, condition or age.

##Classify AD proteins

```{r}

##Prediction for the rest of the proteins


##Filter out training proteins

SampleBtest <- SampleBmerged %>%
  anti_join(down_SampleB) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleBtest$predicted.locations_B <- predict(svM_B, newdata = SampleBtest)
SampleBtest <- SampleBtest %>%
  rownames_to_column()

table(SampleBtest$predicted.locations_B)


```

##Sample D

```{r}
##Filter out training proteins

SampleDtest <- SampleDmerged %>%
  anti_join(down_SampleD) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleDtest$predicted.locations_D <- predict(svM_D, newdata = SampleDtest)
SampleDtest <- SampleDtest %>%
  rownames_to_column()

```

##Sample F

```{r}
##Filter out training proteins

SampleFtest <- SampleFmerged %>%
  anti_join(down_SampleF) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleFtest$predicted.locations_F <- predict(svM_F, newdata = SampleFtest)
SampleFtest <- SampleFtest %>%
  rownames_to_column()

```

##Sample H

```{r}
##Filter out training proteins

SampleHtest <- SampleHmerged %>%
  anti_join(down_SampleH) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleHtest$predicted.locations_H <- predict(svM_H, newdata = SampleHtest)
SampleHtest <- SampleHtest %>%
  rownames_to_column()

```


##Sample J

```{r}
##Filter out training proteins

SampleJtest <- SampleJmerged %>%
  anti_join(down_SampleJ) %>%
  dplyr::select(-Location) %>%
  column_to_rownames(var = "UniprotID")


SampleJtest$predicted.locations_J <- predict(svM_J, newdata = SampleJtest)
SampleJtest <- SampleJtest %>%
  rownames_to_column()

```


##Results table

```{r}
tmpB <- SampleBtest %>%
  dplyr::select(rowname, predicted.locations_B)
tmpD <- SampleDtest %>%
  dplyr::select(rowname, predicted.locations_D)
tmpF <- SampleFtest %>%
  dplyr::select(rowname, predicted.locations_F)
tmpH <- SampleHtest %>%
  dplyr::select(rowname, predicted.locations_H)
tmpJ <- SampleJtest %>%
  dplyr::select(rowname, predicted.locations_J)




test_results <- merge(tmpB, tmpD, by = "rowname", all = TRUE)
test_results <- merge(test_results, tmpF, by = "rowname", all = TRUE)
test_results <- merge(test_results, tmpH, by = "rowname", all = TRUE)
test_results <- merge(test_results, tmpJ, by = "rowname", all = TRUE)

test_results$N <- apply(!is.na(test_results[,2:6]), 1, sum)
n_group <- test_results %>%
  dplyr::select(rowname, N)

tmp <- test_results %>%
  pivot_longer(-c(rowname, N), names_to = "Sample", values_to = "Location") 

tmp <- data.frame(table(tmp$rowname, tmp$Location))
results_consensus <- tmp %>%
  group_by(Var1) %>%
  slice_max(Freq)


results_consensus <- merge(results_consensus, n_group, by.x = "Var1", by.y = "rowname")
results_consensus <- results_consensus %>%
  mutate(prop = (Freq/N)*100)

clear_location <- results_consensus %>%
  filter(prop > 50) %>%
  mutate(Confidence = ifelse(prop > 70, "High", "Medium")) %>%
  dplyr::select(Var1, Var2, Confidence) %>%
  dplyr::rename(Location = Var2)


ties <- results_consensus %>%
  filter(!(Var1 %in% clear_location$Var1)) %>%
  arrange(Var1) 


fourty_percent <- ties %>%
  filter(prop >= 40) %>%
  group_by(Var1) %>%
  mutate(id = row_number()) %>%
  dplyr::select(Var1, id, Var2) %>%
  arrange(Var2) %>%
  pivot_wider(names_from = id, values_from = Var2) %>%
  dplyr::rename(Location_1 = `1`, Location_2 = `2`) %>%
  unite("Split_location", Location_1:Location_2, remove = FALSE) 

fourty_percent$Split_location <- gsub("_NA", "", fourty_percent$Split_location)
low_confidence <- fourty_percent %>%
  mutate(Confidence = "Low") %>%
  dplyr::select(Var1, Split_location, Confidence) %>%
  dplyr::rename(Location = Split_location)

remaining <- results_consensus %>%
  filter(!(Var1 %in% clear_location$Var1)) %>%
  filter(!(Var1 %in% fourty_percent$Var1)) %>%
  mutate(Location = NA) %>%
  mutate(Confidence = "None") %>%
  dplyr::select(Var1, Location, Confidence) %>%
  unique()
 
##Change confidence from high group to high and Medium
AD_final_results <- rbind(clear_location, low_confidence, remaining)

tmp <- data.frame(table(AD_final_results$Location))

table(AD_final_results$Location, AD_final_results$Confidence)

final_results <- final_results %>%
  dplyr::rename(Location_control = Location) %>%
  dplyr::rename(Confidence_control = Confidence)

AD_final_results <- AD_final_results %>%
  dplyr::rename(Location_AD = Location) %>%
  dplyr::rename(Confidence_AD = Confidence)

data <- read.xlsx("Input data/ProteinLevelRawRenamed.xlsx")  
ProteinIDs <- data %>%
  dplyr::select(Protein.IDs) %>%
  separate(Protein.IDs, into = c(NA, "UniprotID", "Uniprot_name"), sep = "\\|")

all_SVM_results <- merge(final_results, AD_final_results, all = TRUE)
all_SVM_results <- merge(all_SVM_results, ProteinIDs, all.y = FALSE, by.x = "Var1", by.y = "UniprotID")
all_SVM_results <- all_SVM_results %>%
  dplyr::rename(Uniprot_ID = Var1) %>%
  dplyr::select(Uniprot_ID, Uniprot_name, Location_control, Confidence_control, Location_AD, Confidence_AD)

write_csv(all_SVM_results, "Supp tables/Table_S5_SVM_results.csv")

```
